---
title: "Project 1"
author: "Olai Gaarn Skogen"
date: '2024-02-10'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
```


# Problem 1A

Correlation functions.

Powered exponential function and Matérn function

```{r}
powered_exp_corr <- function(dist, range, alpha){
  return(exp(-(dist/range)^alpha))
}

matern_corr <- function(dist, range, smooth){
  nu = sqrt(8*smooth)*dist/range
  return(2^(1-smooth) * nu * BesselK(nu) / gamma(smooth))
}

```

Semi-variogram function

## How do I calc semi variogram from covariance

```{r}
semi_variogram_from_cov <- function(cov_vec)
```

# Problem 1B

Because $\boldsymbol{X}$ is a Gaussian Random Field it is by definition distributed as a multivariate normal distribution with $\boldsymbol{\mu} = E[\boldsymbol{X}]$ and $\Sigma = Cov[\boldsymbol{X}, \boldsymbol{X}]$. 

```{r}
calc_spatial_cov_matrix <- function(cov_func, s){
  Sigma = c()
  for(i in 1:length(s)){
    dist_i = abs(s[i] - s)
    print(dist_i)
    covar = cov_func(dist_i)
    Sigma = rbind(Sigma, covar)
  }
  return(Sigma)
}
```

```{r}
n = 50
sigma_sq = 5
range = 10
smoothness = 3
pw_exp_cov_func_11 <- function(dist){sigma_sq*powered_exp_corr(dist, range, smoothness)}

s = seq(1, n)
mu = integer(n)
Sigma = calc_spatial_cov_matrix(pw_exp_cov_func_11, s)

print(Sigma)

X = mvrnorm(mu = mu, Sigma = Sigma)
plot(s, X)
```


# Problem 2A

```{r}

```

